\begin{lemma}
    In any causal diamond $p, q, p \wedge q, p \vee q$ we have $d(p, p \vee q) \leq d(q, p \wedge q)$.
\end{lemma}

\begin{proof}

\begin{align*}
    d(p, p \vee q) &= H(p) + H(p \vee q) - 2H(p \vee (p \vee q))\\
    &= H(p) - H(p \vee q)\\
    &= H(p) + H(p \wedge q) - H(p \wedge q) - H(p \vee q)\\
    \overset{\text{supermod}}&{\leq} H(p \wedge q) - H(q)\\
    &= d(q, p \wedge q)
\end{align*} \phantom{\qedhere}

\end{proof}

\begin{result} \label{res:monotone_future_p}
    There exists no bistochastic matrix $D$ such that $E^- (Dp \parallel q) < E^- (p \parallel q)$.
\end{result}

\begin{proof}
Let $p'$ be a vector majorized by $p$, i.e. there exists a bistochastic matrix $D$ such that $p' = Dp$. Let $q' = p' \vee q$, and let $p'' = p \wedge q'$. By hypothesis and by definition of $q'$, $p'$ is majorized by both $p$ and $q'$, which is equivalent to $p'$ being majorized by $p \wedge q'$. 

\begin{align}
    \implies d(p'', q') = H(p \wedge q') - H(q') \leq H(p') - H(q') = d(p', q') \label{eq:construction}
\end{align}

\begin{align*}
    \implies \min_{p' \prec p} E^- (p' \parallel q) &= \min_{p' \prec p} d(p', p' \vee q)\\
    &= \min_{p' \prec p} d(p', q')\\
    \overset{\text{(\ref{eq:construction})}}&{=} \min_{p' \prec p} d(p'', q')\\
    &= \min_{p' \prec p} d(p \wedge q', q')\\
    \overset{\text{Lemma 3}}&{\geq} \min_{p' \prec p} d(p, p \vee q')\\
    &= \min_{p' \prec p} d(p, p \vee (p' \vee q))\\
    &= \min_{p' \prec p} d(p, (p \vee p') \vee q)\\
    &= d(p, p \vee q)\\
    &= E^- (p \parallel q). \qedhere
\end{align*} 

\end{proof}

\begin{corollary}
    Let $\ket{\psi}$ and $\ket{\phi}$ be two quantum states, and let $\lambda_\psi$ and $\lambda_\phi$ be the associated Schmidt vectors. If $\ket{\psi} \overset{\text{LOCC}}{\longrightarrow} \ket{\phi}$ with probability 1, then $E^- (\lambda_\psi \parallel \lambda_\alpha) \geq E^- (\lambda_\phi \parallel \lambda_\alpha)$ with $\lambda_\alpha$ some probability vector.
\end{corollary}

~

\begin{corollary}
    Let $\rho_{AB}$ be a bipartite quantum state, and let $\rho_A$ and $\rho_B$ be the reduced states. Let $\lambda_{AB}, \lambda_A$ and $\lambda_B$ be the vectors of eigenvalues of their density matrices. Then, if $\rho_{AB}$ is separable, $E^- (\lambda_{AB} \parallel \lambda_B) \geq E^- (\lambda_A \parallel \lambda_B)$ and $E^- (\lambda_{AB} \parallel \lambda_A) \geq E^- (\lambda_B \parallel \lambda_A)$.
\end{corollary}

~

\noindent One would hope that the analogue of result \ref{res:monotone_down} would hold for the other measure of incomparability $E^+$, however it does not. The following counterexample shows that a bistochastic matrix $D$ such that $E^+ (Dp || q) < E^+ (p || q)$ can exist, but also that a bistochastic matrix $D'$ such that $E^+ (D'p || q) > E^+ (p || q)$ can exist too. Let $q = (0.6, 0.4)$, $p = (0.7, 0.29, 0.1)$, $p' = p \wedge q$ and $p'' = (0.7, 0.15, 0.15)$. One can verify that $p$ majorizes both $p'$ and $p''$, yet $E^+ (p || q) = 0.0939$ bits, $E^+ (p' || q) = 0$ bits and $E^+ (p'' || q) = 0.171$ bits.